NN Portfolio Optimierung

### 2024-02-14 paper/Paper.tex
finish methods section images and add val-loss to results

            \textbf{Normal Distribution Density Function  $ X \sim \mathcal{N}(\mu^2, \sigma)  $}

@@

            \textbf{Log-Normal Distribution Density Function  $ X \sim L\mathcal{N}(\mu^2, \sigma) $}

@@

                The next important topic to cover is the loss function used since the learning process can only happen 
                if the model has a loss function to calculate the error. 
                The loss function in the models I built was the cross-entropy loss function. 
                These functions, used in classification problems, calculate how accurate the machine learning 
                or deep learning model is by defining the difference between the estimated probability 
                with the desired outcome \cite{365team2023}. These functions come in two versions, 
                one for binary classification and one for categorical classification. 
                My model uses categorical cross-entropy because it is a multi-class classification problem, 
                which means there is only one correct class per x input or image. Had I evaluated if multiple graphs were in an image,
                then it would become a multi-label classification problem, so the final prediction vector would have
                a dimensionality of the number of classes, but each element in the vector could be 0 or 1 depending 
                if the image contains two classes. For example, if an image had two graphs with one uniform distribution 
                and one normal distribution, then it would look like [0, 0, 1, 1] (classes sorted alphabetically) \cite{gomez2018}.

@@

                        \includegraphics[scale=0.5]{types-of-graphs.png}
                        \caption{Long Short-Term Memory Unit \cite{zhang2020}}

@@

            The implementation of this project wouldn’t be possible without the various tools used. 
            I’ll describe them all briefly and list out the versions used for the sake of reproducibility. 
            TensorFlow \cite{abadi2016-tensorflow} is an end-to-end platform for building machine learning models 
            and that works with Keras \cite{chollet2015-keras}, which is a deep learning API written in Python connected to TensorFlow, 
            to create all of the models in this project. 
            Python comes with a plethora of libraries to make working with data and images easier. 
            First there is NumPy \cite{harris2020-numpy} -- a math library with powerful ndarrays. 
            Then there is pandas \cite{reback2020-pandas} \cite{mckinney2010-pandas} 
            which allows for the creation and manipulation of dataframes, which themselves are made up of NumPy’s arrays. 
            The next important library is Matplotlib \cite{hunter2007-matplotlib} 
            where all of the generated graphs used in the models were plotted. 
            Lastly, one of the most important libraires for image classification using TensorFlow 
            and Keras is Pillow \cite{clark2015pillow} since it allows the user to manipulate individual pixels of an image using Python. 
            The PIL Image is a class that can be used and returned using TensorFlow.

            Table of Versions

@@

                \scalebox{1.2}[1.2]{
                \begin{tabular}{ | c | c | }
                \hline
                \textbf{Library}    &       \textbf{Versionen} \\
                \hline
                python              &       3.10.13 \\ 
                tensorflow          &       2.10.0 \\
                numpy               &       1.26.0 \\
                pandas              &       2.1.1 \\
                matplotlib          &       3.8.0 \\

@@

                \caption{Python und Library Versionen}
                \label{pl-versionen}

@@

            In total, four simple graph classifiers were created -- each with a specific purpose of evaluation in mind. 
            The idea behind them was to create a base assumption that the graphs I generated can be considered a realistic dataset 
            one could get if they scraped from various sources. 
            The main goal is to be able to classify graphs into their probability distribution, 
            but this step acts as a bridge and quality assurance for the generated graph dataset 
            which will be used in the distribution graph classifiers. 
            Each model has a specific name consisting of acronyms to distinguish itself: 
            CIFAR is the natural images, SCP is the scraped graphs, GEN stands for the generated graphs, 
            and 1 indicates the model structure. This results in the four models: CIFAR\_SCP\_1, CIFAR\_GEN\_1, GEN\_SCP\_1, and CIFAR\_GEN\_SCP\_1.

            CIFAR\_SCP\_1

@@

            CIFAR\_GEN\_1

            The second model, much like the first one, answers the question: 
            are the generated graphs distinguishable from the natural images as well? 
            This was the first test for the generated graphs and much like the first model, 
            should have achieved high accuracy. The main question here would be the difference in accuracy between CIFAR\_SCP\_1 
            and CIFAR\_GEN\_1 since a higher accuracy in the former would indicate 
            that my generated graphs have some similarities with natural images that are not apparent in the scraped graphs. 
            This could be turned around and a much higher accuracy in CIFAR\_GEN\_1 could indicate 
            that one of my original theories about the scraped dataset could be correct 
            and that is the ‘graph-like’ nature of the images in the scraped dataset which are more artistic 
            or non-numerical representations of graphs.

            GEN\_SCP\_1

            This was a major test of the simple graph classifier 
            since it will have indicated if my generated graphs were randomized enough to simulate what a bunch of scraped graphs would be. 
            The desired outcome here would be a lower accuracy of around 50 percent to show the interchangeability of the two datasets. 
            A high accuracy would indicate the need to randomize the generated graphs more and put them through more preprocessing steps.

            CIFAR\_GEN\_SCP\_1

            Lastly, all of the datasets were used in a model to classify them as natural, generated, or scraped. 
            The outcome of this would not be as important as the first three, 
            but it would have reinforced the previous intermediate conclusions about the nature of the datasets. 
            A very high accuracy would have meant all datasets are highly distinguishable from one another 
            and point towards tweaking the generated graphs.

            \newpage
            \begin{Large} \[ V_p = \sum_{j=1}^{A}\sum_{i=1}^{M}P^T_{i+1} \cdot w_i \] \end{Large}

@@

            NOTE: \newline
            \url{https://github.com/p-spohr/NN-Graph-Classifier/tree/main/classifiers/distribution_graph_classifiers}

            The images used, total number of images, training images, validation images, 
            and model structure are all the same, except for the dimensions of the image. 
            The reason for this is how much information is lost when the image is 32x32. 
            It was easy to assume that the model would perform poorly when the density function line almost completely disappears. 
            My solution was to increase the dimensions and thereby its quality. 
            The classes, y target, are as previously stated: exp, lognorm, norm, and unif.

@@

@@

@@

                    \includegraphics[scale=0.5]{types-of-graphs.png}
                    \caption{Summe monatlichen Renditen von 1990 bis 2020}
                    \label{s-m-renditen-fig}

@@

                    \begin{tabular}{ | l | l | }

                        \hline
                        \textbf{Portfolio}   & \textbf{Summe Monatlichen Renditen \%} \\
                        \hline
                        OPT mit RFR          &      6604,73 \\            
                        OPT ohne RFR         &      5331,83 \\
                        EQU                  &      213,32 \\                 
                        GSPC                 &      248,56 \\

                        \hline

                    \end{tabular}
                    \caption{Summe Monatlichen Renditen}
                    \label{s-m-renditen-tab}
                \end{center}
            \end{table}

@@

                \begin{tabular}{ | l | l | }

                    \hline
                    \textbf{Portfolio}   & \textbf{Monatlicher Durchschnitt} \\
                    \hline
                    OPT mit RFR          & 39,9735  \\          
                    OPT ohne RFR         & 41,1696 \\
                    EQU                  & 52,8962  \\              
                    GSPC                 & 72,302  \\

                    \hline

                \end{tabular}
                \caption{Gesamtsumme der Sharpe-Quotienten ohne den Risikofreinen Zinssatz}
                \label{gs-sq-ohne-rfz}

                \begin{tabular}{ | l | l | }

                    \hline
                    \textbf{Portfolio}   & \textbf{Monatlicher Durchschnitt} \\
                    \hline
                    OPT mit RFR          & -0,0026 \\      
                    OPT ohne RFR         & 0,0225 \\
                    EQU                  & 0,0934 \\            
                    GSPC                 & 1,7403 \\

                    \hline

                \end{tabular}
                \caption{Monatlicher Durchschnitt der Sharpe-Quotienten ohne den Risikofreinen Zinssatz}
                \label{md-sq-ohne-rfz}

                \begin{tabular}{ | l | l | }

                    \hline
                    \textbf{Portfolio}   & \textbf{Monatlicher Durchschnitt} \\
                    \hline
                    OPT mit RFR          & 30,2492  \\          
                    OPT ohne RFR         & 28,7633 \\
                    EQU                  & -345,5259  \\              
                    GSPC                 & -395,0716  \\

                    \hline

                \end{tabular}
                \caption{Gesamtsumme der Sharpe-Quotienten mit dem Risikofreinen Zinssatz}
                \label{gs-sq-mit-rfz}

                \begin{tabular}{ | l | l | }

                    \hline
                    \textbf{Portfolio}   & \textbf{Monatlicher Durchschnitt} \\
                    \hline
                    OPT mit RFR          & 0,08449 \\      
                    OPT ohne RFR         & 0,08034 \\
                    EQU                  & -0,9651 \\            
                    GSPC                 & -1,1035 \\

                    \hline

                \end{tabular}
                \caption{Monatlicher Durchschnitt der Sharpe-Quotienten mit dem Risikofreinen Zinssatz}
                \label{md-sq-mit-rfz}

@@

            As a whole, I can say that the overall goals were accomplished. 
            I have a model that can successfully identify graphs from natural images 
            with greater than 98 percent accuracy and another model that can successfully identify graphs 
            with four different distribution types with around 93 percent accuracy. 
            Addendum to these results is the caveat from earlier that the generated graphs could be considered as a disjunct subset, 
            if one considers the scraped graphs to be a genuine sample of possible and currently available graphs. 
            The GEN\_SCP\_1’s high accuracy points towards this theory but a large limiting factor is the small sample size -- 
            only 8,000 generated graphs and 7,753 scraped graphs. 
            Some results that surprised me was that scraped graphs in CIFAR\_GEN\_1 were misclassified as natural images 
            around 63 percent of the time. This could indicate 
            that the generated graphs don’t consider the more chaotic nature of graph design. 
            It would be safe to say that the generated graphs are very conservative in appearance. 
            However, this was an original design choice to reduce the amount of variables in the graphs design, 
            like not having two graphs in one image or keeping to the standard of having a complete graph in each image. 
            Many graphs, like the ones in the scraped dataset 
            and the ones I used in the final test evaluation don’t have any axes, labels, or titles. 
            This enforces the idea that many of the images in SCP and many graphs in general are more ‘graph-like’ 
            and not graphs in a very defined sense.
            Another surprise was the difference in how the uniform distribution graphs were getting misclassified.
            Uniform distribution graphs in DIST\_153x115\_1 were classified accurately around 99 percent of the time, 
            but 20 percent of uniform graphs in CIFAR\_SCP\_1 were misclassified as natural images. 
            Why were the graphs easily classified in DIST\_153x115\_1 but not in CIFAR\_SCP\_1? 
            Perhaps graphs with the uniform distribution are unique from the other distribution graphs but not very unique from natural images.

@@

>>>>

            \textbf{Normal Distribution Density Function  $ X \sim \mathcal{N}(\mu, \sigma^2)  $}

@@

            \textbf{Log-Normal Distribution Density Function  $ X \sim L\mathcal{N}(\mu, \sigma^2) $}

@@

                The next important topic to cover is the loss function used since the learning process can only happen 
                if the model has a loss function to calculate the error. 
                The loss function in the models I built was the cross-entropy loss function. 
                These functions, used in classification problems, calculate how accurate the machine learning 
                or deep learning model is by defining the difference between the estimated probability 
                with the desired outcome \cite{365team2023}. These functions come in two versions, 
                one for binary classification and one for categorical classification. 
                My model uses categorical cross-entropy because it is a multi-class classification problem, 
                which means there is only one correct class per x input or image. Had I evaluated if multiple graphs were in an image,
                then it would become a multi-label classification problem, so the final prediction vector would have
                a dimensionality of the number of classes, but each element in the vector could be 0 or 1 depending 
                if the image contains two classes \cite{gomez2018}. For example, if an image had two graphs with one uniform distribution 
                and one normal distribution, then it would look like [0, 0, 1, 1] 
                (classes sorted alphabetically: exp, lognorm, norm, unif).

@@

                        \includegraphics[scale=0.9]{adam-opt.png}
                        \caption{Adam Optimizer Algorithm \cite{kingma2015}}

@@

            The implementation of this project wouldn’t be possible without the various tools used. 
            I’ll describe them all briefly and list out the versions used for the sake of reproducibility in Table \ref{py-versions}. 
            TensorFlow \cite{abadi2016-tensorflow} is an end-to-end platform for building machine learning models 
            and that works with Keras \cite{chollet2015-keras}, which is a deep learning API written in Python connected to TensorFlow, 
            to create all of the models in this project. 
            Python comes with a plethora of libraries to make working with data and images easier. 
            First there is NumPy \cite{harris2020-numpy} -- a math library with powerful ndarrays. 
            Then there is pandas \cite{reback2020-pandas} \cite{mckinney2010-pandas} 
            which allows for the creation and manipulation of dataframes, which themselves are made up of NumPy’s arrays. 
            The next important library is Matplotlib \cite{hunter2007-matplotlib} 
            where all of the generated graphs used in the models were plotted. 
            Lastly, one of the most important libraires for image classification using TensorFlow 
            and Keras is Pillow \cite{clark2015pillow} since it allows the user to manipulate individual pixels of an image using Python. 
            The PIL Image is a class that can be used and returned using TensorFlow.

@@

                \scalebox{1.2}[1.2]{
                \begin{tabular}{ | c | c | }
                \hline
                \textbf{Library}    &       \textbf{Versions} \\
                \hline
                python              &       3.10.13 \\ 
                tensorflow          &       2.10.0 \\
                numpy               &       1.26.0 \\
                pandas              &       2.1.1 \\
                matplotlib          &       3.8.0 \\
                pillow              &       10.0.1 \\

@@

                \caption{Python und Library Versions}
                \label{py-versions}

@@

            In total, four simple graph classifiers were created -- each with a specific purpose of evaluation in mind. 
            The idea behind them was to create a base assumption that the graphs I generated can be considered a realistic dataset 
            one could get if they scraped from various sources. 
            The main goal is to be able to classify graphs into their probability distribution, 
            but this step acts as a bridge and quality assurance for the generated graph dataset 
            which will be used in the distribution graph classifiers. 
            Each model has a specific name consisting of acronyms to distinguish itself: 
            CIFAR is the natural images, SCP is the scraped graphs, GEN stands for the generated graphs, 
            and 1 indicates the model structure. This results in the four models: CIFAR\_SCP\_1, CIFAR\_GEN\_1, GEN\_SCP\_1, 
            and CIFAR\_GEN\_SCP\_1. The model struture for all classifiers is shown by layer in Table \ref{model-layer-info}.

                \scalebox{0.9}{
                \begin{tabular}{ | l | l | l | }
                \hline
                \textbf{Layer} &  \textbf{Type}     &  \textbf{Parameters}    \\
                \hline
                1              &  Rescaling         &  (1.0/255, input\_shape = (H, W, 3))                \\ 
                2              &  Conv2D            &  (filter\_count = 16, kernal\_size = 3, padding='same', activation = 'relu')   \\
                3              &  MaxPooling2D      &  none                        \\
                4              &  Conv2D            &  (32, 3, 'same', 'relu')                        \\
                5              &  MaxPooling2D      &  none                       \\
                6              &  Conv2D            &  (64, 3, 'same', 'relu')                           \\
                7              &  MaxPooling2D      &  none                           \\
                8              &  Flatten           &  none                            \\
                9              &  Dense             &  (units = 128, 'relu')                            \\
                10             &  Dense             &  (number\_of\_classes = 4)                            \\

                \caption{Model 1 Layer Information}
                \label{model-layer-info}

                \scalebox{1.2}{
                \begin{tabular}{ | c | c | c | }
                \hline
                \textbf{Model Info}    &       \textbf{CIFAR\_SCP\_1}     &       \textbf{CIFAR\_GEN\_1} \\
                \hline
                x                      &       32x32 images               &       32x32 images \\ 
                y                      &       [graph, natural]           &       [graph, natural] \\
                total                  &       17,753                     &       18,000 \\
                train                  &       14,203                     &       14,400\\
                val                    &       3,550                      &       3600 \\
                model                  &       1                          &       1 \\

                \caption{Model Information for classifying graphs from natural images}
                \label{cifar-graphs-1}

@@

            The second model, much like the first one, answers the question: 
            are the generated graphs distinguishable from the natural images as well? 
            This was the first test for the generated graphs and much like the first model, 
            should have achieved high accuracy. The main question here would be the difference in accuracy between CIFAR\_SCP\_1 
            and CIFAR\_GEN\_1 since a higher accuracy in the former would indicate 
            that my generated graphs have some similarities with natural images that are not apparent in the scraped graphs. 
            This could be turned around and a much higher accuracy in CIFAR\_GEN\_1 could indicate 
            that one of my original theories about the scraped dataset could be correct 
            and that is the ‘graph-like’ nature of the images in the scraped dataset which are more artistic 
            or non-numerical representations of graphs. 
            In Table \ref{cifar-graphs-1}, one can see that the input in CIFAR\_SCP\_1 and CIFAR\_GEN\_1 are balanced.

            GEN\_SCP\_1 was a major test of the simple graph classifier 
            since it will have indicated if my generated graphs were randomized enough to simulate what a bunch of scraped graphs would be.
            The desired outcome here would be a lower accuracy of around 50 percent to show the interchangeability of the two datasets. 
            A high accuracy would indicate the need to randomize the generated graphs more and put them through more preprocessing steps.

            Lastly, all of the datasets were used in a model to classify them as natural, generated, or scraped. 
            The outcome of this would not be as important as the first three, 
            but it would have reinforced the previous intermediate conclusions about the nature of the datasets. 
            A very high accuracy would have meant all datasets are highly distinguishable from one another 
            and point towards tweaking the generated graphs. The information for GEN\_SCP\_1 and CIFAR\_GEN\_SCP\_1
            can be seen in Table \ref{cifar-gen-scp-1}.

                \scalebox{1.2}{
                \begin{tabular}{ | c | c | c | }
                \hline
                \textbf{Model Info}    &       \textbf{GEN\_SCP\_1}       &       \textbf{CIFAR\_GEN\_SCP\_1} \\
                \hline
                x                      &       32x32 images               &       32x32 images \\ 
                y                      &       [generated, scraped]       &       [generated, natural, scraped] \\
                total                  &       15,753                     &       25,753 \\
                train                  &       12,603                     &       20,603\\
                val                    &       3,150                      &       5,150 \\
                model                  &       1                          &       1 \\

                \caption{Model Information for classifying generated from scraped graphs}
                \label{cifar-gen-scp-1}

@@

                \scalebox{0.9}{
                \begin{tabular}{ | c | c | c | c | }
                \hline
                \textbf{Model Info} &  \textbf{DIST\_32x32\_1}      &  \textbf{DIST\_115x86\_1}    &  \textbf{DIST\_153x115\_1} \\
                \hline
                x                   &  32x32 images                 &  115x86 images               &  153x115 images \\ 
                y                   &  [exp, lognorm, norm, unif]   &  [exp, lognorm, norm, unif]  &  [exp, lognorm, norm, unif] \\
                total               &  8,000                        &  8,000                       &  8,000 \\
                train               &  6,400                        &  6,400                       &  6,400 \\
                val                 &  1,600                        &  1,600                       &  1,600 \\
                model               &  1                            &  1                           &  1 \\

                \caption{Model Information for all distribution graph classifiers}
                \label{dist-dims-1}

            The images used, total number of images, training images, validation images, 
            and model structure are all the same, except for the dimensions of the image and that can be seen in Table \ref{dist-dims-1}.
            The reason for this is how much information is lost when the image is 32x32. 
            It was easy to assume that the model would perform poorly when the density function line almost completely disappears. 
            My solution was to increase the dimensions and thereby its quality. 
            The classes, y target, are as previously stated: exp, lognorm, norm, and unif.

@@

            \begin{figure}

                    \includegraphics[scale=0.8]{CIFAR_SCP_1_HIST_RESULTS.png}
                    \caption{CIFAR\_SCP\_1 Training and Validation Accuracy and Loss}
                    \label{cifar-scp-acc-loss}

                    \includegraphics[scale=0.8]{CIFAR_GEN_1_HIST_RESULTS.png}
                    \caption{CIFAR\_GEN\_1 Training and Validation Accuracy and Loss}
                    \label{cifar-gen-acc-loss}

            \begin{figure}

                    \includegraphics[scale=0.8]{GEN_SCP_1_HIST_RESULTS.png}
                    \caption{GEN\_SCP\_1 Training and Validation Accuracy and Loss}
                    \label{gen-scp-acc-loss}

                    \includegraphics[scale=0.8]{CIFAR_GEN_SCP_1_HIST_RESULTS.png}
                    \caption{GEN\_SCP\_1 Training and Validation Accuracy and Loss}
                    \label{cifar-gen-scp-acc-loss}

@@

            \begin{figure}

@@

                    \includegraphics[scale=0.6]{DIST_32x32_1_HIST_RESULTS.png}
                    \caption{DIST\_32x32\_1 Training and Validation Accuracy and Loss}
                    \label{dist-32-32-val-loss}

                    \includegraphics[scale=0.6]{DIST_115x86_1_HIST_RESULTS.png}
                    \caption{DIST\_115x86\_1 Training and Validation Accuracy and Loss}
                    \label{dist-115-86-val-loss}

                    \includegraphics[scale=0.6]{DIST_153x115_1_HIST_RESULTS.png}
                    \caption{DIST\_153x115\_1 Training and Validation Accuracy and Loss}
                    \label{dist-153-115-val-loss}

@@

@@

@@

            As a whole, I can say that the overall goals were accomplished. 
            I have a model that can successfully identify graphs from natural images 
            with greater than 98 percent accuracy and another model that can successfully identify graphs 
            with four different distribution types with around 93 percent accuracy. 
            Addendum to these results is the caveat from earlier that the generated graphs could be considered as a disjunct subset, 
            if one considers the scraped graphs to be a genuine sample of possible and currently available graphs. 
            The GEN\_SCP\_1’s high accuracy points towards this theory but a large limiting factor is the small sample size -- 
            only 8,000 generated graphs and 7,753 scraped graphs. 
            Some results that surprised me were that scraped graphs in CIFAR\_GEN\_1 were misclassified as natural images 
            around 63 percent of the time. This could indicate 
            that the generated graphs don’t consider the more chaotic nature of graph design. 
            It would be safe to say that the generated graphs are very conservative in appearance. 
            However, this was an original design choice to reduce the amount of variables in the graphs design, 
            like not having two graphs in one image or keeping to the standard of having a complete graph in each image. 
            Many graphs, like the ones in the scraped dataset 
            and the ones I used in the final test evaluation don’t have any axes, labels, or titles. 
            This enforces the idea that many of the images in SCP and many graphs in general are more ‘graph-like’ 
            and not graphs in a very defined sense.
            Another surprise was the difference in how the uniform distribution graphs were getting misclassified.
            Uniform distribution graphs in DIST\_153x115\_1 were classified accurately around 99 percent of the time, 
            but 20 percent of uniform graphs in CIFAR\_SCP\_1 were misclassified as natural images. 
            Why were the graphs easily classified in DIST\_153x115\_1 but not in CIFAR\_SCP\_1? 
            Perhaps graphs with the uniform distribution are unique from the other distribution graphs but not very unique from natural images.

@@

    \begin{Large} \[ V_p = \sum_{j=1}^{A}\sum_{i=1}^{M}P^T_{i+1} \cdot w_i \] \end{Large}
